---
title: "Assignment 1"
author: "pk673@scarletmail.rutgers.edu"
output: html_document
date: "2023-02-01"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
```{r}
## Assignment : As the professor mentioned, one homework is to create 2 separate matrices: 1 for survived sparrows and 1 for non-survived sparrows. Then, we can calculate col_means, create covariance matrices and correlation matrices.

#install required packages
library(readr)
library(HSAUR2)
library(MVA)


#Import Sparrow dataset 
Bumpus_sparrows <- read_csv("~/Downloads/Bumpus_sparrows.csv")
Bumpus_sparrows
sparr <- Bumpus_sparrows
sparr

#Creating a subset for Survived sparrows(S)
Survived <- subset(sparr, Survivorship == "S")
Survived
is.vector(Survived)
is.matrix(Survived)
class(Survived)
attributes(Survived)

sparr_s <- Survived
sparr_s
sparr_s[,-1]
#We will drop column 1 as it does not contain a number
sparr_s.num <- sparr_s[,-1]
sparr_s.num
# Computing the means of each variable in data frame sparr_s.num
colMeans(sparr_s.num)
# Covariance matrix
cov(sparr_s.num)
# Correlation matrix
cor(sparr_s.num)

#Creating a subset for sparrows that did Not Survive (NS)
NotSurvived <- subset(sparr, Survivorship == "NS")
NotSurvived
is.vector(NotSurvived)
is.matrix(NotSurvived)
class(NotSurvived)
attributes(NotSurvived)

sparr_ns <- NotSurvived
sparr_ns
sparr_ns[,-1]
#We will drop column 1 as it does not contain a number
sparr_ns.num <- sparr_s[,-1]
sparr_ns.num
# Computing the means of each variable in data frame sparr_ns.num
colMeans(sparr_ns.num)
# Covariance matrix
cov(sparr_ns.num)
# Correlation matrix
cor(sparr_ns.num)
