---
title: "Assignment 2"
author: "pk673@scarletmail.rutgers.edu"
output: html_document
date: "2023-02-11"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE)
```

## R Markdown
```{r}
#install required packages
library(readr)
library(SciViews)
library(scatterplot3d)
library(car)
library(lattice)
library(GGally)
library(ggthemes)
library(dplyr)
library(tidyverse)
library(grid)
library(gridExtra)
library(RColorBrewer)
library(HSAUR2)
library(MVA)
library(ggridges)
library(cowplot)
library(hexbin)

#Import Sparrow dataset 
Bumpus_sparrows <- read_csv("~/Downloads/Bumpus_sparrows.csv")
Bumpus_sparrows
sparrows <- Bumpus_sparrows
sparrows

boxplot(sparrows[,2:6])
stars(sparrows,labels = sparrows$Survivorship) # for understanding if observations are related or not, if they are similar

# box plot
ggplot(sparrows, aes(x=Survivorship, y=Total_length)) + geom_boxplot()
## Sparrows with the higher total length did not survive as compared to sparrows with smaller total length


attach(sparrows)

#Plots
sparr.birds <- data.frame(as.numeric(rownames(sparrows)),sparrows[,2:6])
labs.diagonal <- c("Bird","Total length","Alar extent","L. beak & head","L. humerous","L. keel & sternum")

plot(Total_length, Alar_extent,xlab="Total Length (mm)",ylab="Alar extent (nm)",pch=c(16,1))

# Correlations
pairs(sparrows[,2:6])

# A Different View on Correlation Matrix
library(SciViews)
pairs(sparr.birds, diag.panel = panel.boxplot, labels=labs.diagonal,pch=c(1,16),font.labels=2)

# 3 Dimensional Plots
library(scatterplot3d)
s3d <- scatterplot3d(Alar_extent,Total_length,L_beak_head,pch=c(1,16)[as.numeric(Survivorship)],xlab="Alar extent", ylab="", angle=45,zlab="Length of beak and head", lty.hide=2,type="h",y.margin.add=0.1,font.axis=2,font.lab=2)

legend(s3d$xyz.convert(238, 160, 34.1),c("Non-survivor","Survivor"),pch=c(1,16),text.font=2)

#scatterplotMatrix
library(car)
scatterplotMatrix(~Total_length+Alar_extent+L_beak_head+L_humerous+L_keel_sternum | Survivorship, data=sparr.birds, var.labels=labs.diagonal,cex.labels=0.7, diagonal="boxplot",smooth=FALSE,reg.line=FALSE,pch=c(1,16),col=rep("black",2), legend.plot=FALSE)

library(lattice)
super.sym <- trellis.par.get("superpose.symbol")
super.sym$superpose.symbol$pch <- c(1,16,rep(1,5))
super.sym$superpose.symbol$col <- c("#0000FF","#FF69B4",rep(1,5))
trellis.par.set(super.sym)
splom(~sparr.birds, groups = Survivorship, data = sparr.birds, ps=0.5, varname.cex = .5,panel = panel.superpose,key = list(columns = 2,points = list(pch = super.sym$pch[1:2], col=super.sym$col[1:2]),text = list(c("Non-survivor", "Survivor"))))

library(GGally)
ggscatmat(sparrows, columns=2:6, color="Survivorship")


# bar chart
ggplot(sparrows, aes(x=L_keel_sternum) )+ geom_bar(position="stack",fill='blue') 
## Most number of sparrows have the length of keel_sternum equal to 20.3

# histogram
ggplot(sparrows, aes(x=Total_length))+geom_histogram(fill='red', color='black', binwidth=5)
## 20 sparrows have the Total_length between 152.5 to 157.5 while 2 sparrows have it in the range 147.5 to 152.5


#exploring GGplot

p <- ggplot(data = sparrows, mapping = aes(x = Total_length,y = L_beak_head))
p + geom_point() 
p+ geom_point() + geom_smooth()
p + geom_point() + geom_smooth(method = "lm") 
p + geom_point() + geom_smooth(method = "gam") + scale_x_log10()

p + geom_point() +geom_smooth(method = "gam") + scale_x_log10(labels = scales::dollar)

p <-  ggplot(data = sparrows, mapping = aes(x = Total_length,y = L_beak_head, color = Survivorship))

p + geom_point(color = "purple") + geom_smooth(method = "loess") + scale_x_log10()


p + geom_point(alpha = 0.3) + geom_smooth(color = "orange", se = FALSE, size = 8, method = "lm") + scale_x_log10()

p + geom_point(alpha = 0.3) + geom_smooth(method = "gam") + scale_x_log10(labels = scales::dollar) + labs(x = "Total length", y = "Beak Head",
         title = "Survivorship of Sparrows",
         subtitle = "Survivorship vs non-Survivorship",
         caption = "Source: Bumpus_sparrows")

p <- ggplot(data = sparrows, mapping = aes(x = Total_length,y = L_beak_head, color = Survivorship))
p + geom_point() + geom_smooth(method = "loess") + scale_x_log10()


p <- ggplot(data = sparrows, mapping = aes(x = Total_length,y = L_beak_head, color = Survivorship, fill = Survivorship))
p + geom_point() + geom_smooth(method = "loess") + scale_x_log10()


p <- ggplot(data = sparrows, mapping = aes(x = Total_length,y = L_beak_head))
p + geom_point(mapping = aes(color = Survivorship)) +
    geom_smooth(method = "loess") +
    scale_x_log10()

ggplot(sparrows, aes(x=Alar_extent,y=Total_length)) + facet_wrap(~Survivorship) + geom_point()



# violin plot 
ggplot(sparrows, aes(x=Total_length, y=Alar_extent)) + geom_violin()


# hexbin
#ggplot(sparrows, aes(x=Total_length, y=Alar_extent)) + geom_hex() 
ggplot(data = sparrows, mapping = aes(x = Total_length,y = L_beak_head))+ geom_hex() 


# density plot and ggridges
ggplot(sparrows, aes(x=L_humerous)) + geom_density() 


ggplot(sparrows, aes(x=Total_length, y=Survivorship)) + geom_density_ridges() 


# with ggthemes  

lastplot <- ggplot(data = sparrows, mapping = aes(x = Total_length,y = L_beak_head))
lastplot <- lastplot + geom_point(mapping = aes(color = Survivorship)) + stat_smooth() + 
  labs(x="Total length of sparrows", y="Beak head of sparrows", title="Sparrows Data") 

lastplot + theme_bw()
lastplot + theme_cowplot()
lastplot + theme_dark()
lastplot + theme_economist()
lastplot + theme_fivethirtyeight()
lastplot + theme_tufte()
lastplot + theme_wsj()


